\documentclass[11pt]{article}

\usepackage{fullpage} 
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{pgf}
\usepackage{tikz}
\usetikzlibrary{arrows,automata}

\usepackage{indentfirst}

\newcommand{\question}[2] {\vspace{0.3in}\noindent{\subsection*{Question #1. #2} \vspace{0.15in}}}

\renewcommand{\part}[1] {{\vspace{0.15in}\noindent\textbf (#1)} \vspace{0.10in}}



%  ----------------------------------------------------------------
%                         Start here
% ----------------------------------------------------------------
 
\begin{document}

\title{Assignment \#3} %Replace X with the appropriate number
\author{\Large Gustavo Estrela de Matos\\ %Replace with your name
CSCE 433: Formal Languages and Automata} %If necessary, replace with your course number and title
\date{\today} 

\maketitle

\question{1}{Find the smallest string in $\{a, b\}^{*}$ not in the language corresponding to the given regular expressions. \\
\part{a} 
$b^{*} (ab)^{*} a^{*}$ \\
\part{b}
$(a^* + b^*)(a^* + b^*)(a^* + b^*)$ \\
\part{c} 
$a^*(baa^*)^*b^*$
}

\part{a}
\par This regular expression accepts any string with lenght less or equal to $2$.
\par To make a string that is not accepted we can put a $b$ in the end of the string, considering the fact that this string can't be of format $b^*(ab)^*$ because it would be accepted. The string $abb$, for instance, is not accepted and it is minimal because any string with size less then $abb$ is accepted.

\part{b} 
\par We can see that this regular expression accepts any string with length less or equal to $3$. Also, the look of a string accepted by this language is one that has $3$ or less sequences of $a$'s and $b$'s; if we have more than $3$ sequences the string won't be accepted. Therefore, the string $abab$ is not accepted and is minimal since any string with length less than $4$ is accepted.

\part{c} 
\par First let's observe that any string with size less or equal to $2$ is accepted by this language.
\par To find a string not accepted by this regular expression we are going to exploit the fact that the term in the middle ($baa*$) repeats the string $ba$ everytime that it is repeated.
\par Then, if we have a string that starts with $bb$ the first ($a^*$) and second ($(baa*)^*$) sequences are not in this string, therefore this string should have just $b$'s, then $bba$ is not accepted by the language and it is minimal since every string with size less then $3$ is accepted.

\question{2}{Find a regular expression corresponding to each of the following subsets of $\{a, b\}^*$.\\
\part{a} The language of all strings that begin or end with $aa$ or $bb$. \\
\part{b} The language of all strings with at most two occurrences of the substring $ab$. \\
\part{c} The language of all strings not containing the substring $bba$.}

\part{a}
\par Let's split this subset in simpler subsets:
\begin{itemize}
    \item the subset of strings that start with aa or bb: $(aa + bb)(a + b)^*$
    \item the subset of strings that end with aa or bb: $(a + b)^*(aa + bb)$
\end{itemize}
\par Therefore the subset we want is represented by the regular expression: $(aa + bb)(a + b)^* + (a + b)^*(aa + bb)$

\part{b} 
\par We can think of this subset as "sequences of $a$'s and $b$'s (any string in $\{a, b\}^*$ is a sequence of $a$'s and $b$'s) that changes from $a$'s to $b$'s at most twice". Then the regular expression should look like $a^*b^*a^*b^*$ and since we could have strings starting with $b$ and ending with $a$ without creating any more $ab$ we have that the regular expression is $b^*a^*b^*a^*b^*a^*$.

\part{c}
\par A string that does not contain $bba$ is a string that has a sequence of $a$'s and after a $b$ appears we have three options following this $b$:
\begin{itemize}
    \item a sequence of $a$'s until the end of the string
    \item a sequence of $b$'s until the end of the string
    \item another string that does not contain $bba$
\end{itemize}
\par Which allow us to see that the regular expression for this set is $(a^*b)^*(a* + b^*)$

\question{3}{Use the pumping lemma to prove that the following languages are not regular. \\
\part{a} $L = \{a^nb^nc^{2n} $ $|$ $ n \geq 0\}$ \\
\part{b} $L = \{w $ $|$ $ w \in \{0, 1\}^* $ and $w$ is a palindrome$\}$ \\
\part{c} $L = \{0^i1^j $ $|$ $ j = i$ or $j = 2i\}$
}

\part{a} 
\par Assume that $L$ is a regular language and let $p$ be the pumping constant for $L$.
\par Take the string $s = a^pb^pc^{2p}$. Since $s \in L$ we know that, according to the pumping lemma, $s$ can be written as $s = xyz$ where $y \neq \epsilon$, $|xy| \leq p$ and $xy^iz \in L$ for all integer $i \geq 0$. 
\par Since $|xy| \leq p$ we have that $y = a^j$ for some integer $j > 0$ ($y \neq \epsilon$) such that $s = a^ia^ja^kb^pc^{2p}$ with $i + j + k = p$. Then, using the pumping lemma we know that $s' = xy^0z$ should be in $L$, but $xy^0z = a^{i + k}b^pc^{2p}$ and since $j > 0$ then $i + k \neq p$  and $s' \notin L$ which is a contradiction.
\par Assuming that $L$ is a regular language led us to a contradiction, therefore $L$ is not regular.

\part{b}
\par Assume that $L$ is a regular language and let $p$ be the pumping constant for $L$.
\par Take the string $s = 0^p1^p1^p0^p$. Since $s \in L$ we know that, according to the pumping lemma, $s$ can be written as $s = xyz$ where $y \neq \epsilon$, $|xy| \leq p$ and $xy^iz \in L$ for all integer $i \geq 0$.
\par Since $|xy| \leq p$ we have that $y = 0^j$ for some integer $j > 0$ ($y \neq \epsilon$) such that $s = 0^i0^j0^k1^p1^p0^p$ with $i + j + k = p$. Then, using the pumping lemma we know that $s' = xy^0z$ should be in $L$, but $xy^0z = 0^{i + k}1^p1^p0^p$ and since $j > 0$ then $i + k \neq p$, so $s'$ is not palindrome and  $s' \notin L$ which is a contradiction.
\par Assuming that $L$ is a regular language led us to a contradiction, therefore $L$ is not regular.

\part{c}
\par Assume that $L$ is a regular language and let $p$ be the pumping constant for $L$.
\par Take the string $s = 0^p1^{2p}$. Since $s \in L$ we know that, according to the pumping lemma, $s$ can be written as $s = xyz$ where $y \neq \epsilon$, $|xy| \leq p$ and $xy^iz \in L$ for all integer $i \geq 0$.
\par Since $|xy| \leq p$ we have that $y = 0^j$ for some integer $j > 0$ ($y \neq \epsilon$) such that $s = 0^i0^j0^k1^{2p}$ with $i + j + k = p$. Then, using the pumping lemma we know that $s' = xy^0z$ should be in $L$, but $xy^0z = 0^{i + k}1^{2p}$ and since $j > 0$ we know that $i + k < p$ therefore $i + k \neq p$ and $i + k \neq 2p$, then $s' \notin L$ which is a contradiction.
\par Assuming that $L$ is a regular language led us to a contradiction, therefore $L$ is not regular.

\end{document}
