\documentclass[11pt, a4paper]{awesome-cv}
\usepackage{indentfirst}
% Specify the location of the included fonts
\fontdir[fonts/]

% Color for highlights
% Awesome Colors: awesome-emerald, awesome-skyblue, awesome-red, awesome-pink, awesome-orange
%                 awesome-nephritis, awesome-concrete, awesome-darknight
\colorlet{awesome}{awesome-darknight}
% Uncomment if you would like to specify your own color
% \definecolor{awesome}{HTML}{CA63A8}

% Colors for text
% Uncomment if you would like to specify your own color
% \definecolor{darktext}{HTML}{414141}
% \definecolor{text}{HTML}{333333}
% \definecolor{graytext}{HTML}{5D5D5D}
% \definecolor{lighttext}{HTML}{999999}

% Set false if you don't want to highlight section with awesome color
\setbool{acvSectionColorHighlight}{false}

% If you would like to change the social information separator from a pipe (|) to something else
\renewcommand{\acvHeaderSocialSep}{\quad\textbar\quad}


%-------------------------------------------------------------------------------
%	PERSONAL INFORMATION
%	Comment any of the lines below if they are not required
%-------------------------------------------------------------------------------
\name{Gustavo Estrela de Matos}{}
\address{1100 Hensel Drive, Apt. N302, College Station - Texas}
%\position{Computer Science Student}
\mobile{(+1) 979-402-9614} 
\email{estrela.gustavo.matos@gmail.com}
\github{gustavoem}


%-------------------------------------------------------------------------------
%	LETTER INFORMATION
%	All of the below lines must be filled out
%-------------------------------------------------------------------------------
% The company being applied to
%\recipient
%  {School of Engineering \& Applied Sciences}
%  {Google Inc.\\1600 Amphitheatre Parkway\\Mountain View, CA 94043}
% The date on the letter, default is the date of compilation
%\letterdate{\today}
% The title of the letter
%\lettertitle{Job Application for Software Engineer}
% How the letter is opened
%\letteropening{Dear Mr./Ms./Dr. LastName,}
% How the letter is closed
%\letterclosing{Sincerely,}
% Any enclosures with the letter
%\letterenclosure[Attached]{Curriculum Vitae}


%-------------------------------------------------------------------------------
\begin{document}

% Print the header with above personal informations
\makecvheader

% Print the footer with 3 arguments(<left>, <center>, <right>)
% Leave any of these blank if they are not needed
\makecvfooter
  {\today}
  {}
  {}


% Print the title with above letter informations
%\makelettertitle

%-------------------------------------------------------------------------------
%	LETTER CONTENT
%-------------------------------------------------------------------------------
\begin{cvletter}
\lettersection{About Me}
\par I'm a computer science student from University of S達o Paulo, in Brazil. In the year of 2014 I was selected to participate in the program Science Without Borders, provided by Government of Brazil, and in  August 2015 I moved to College Station, Texas, to study at Texas A\&M University for two semesters.
\par I started my computer science program in University of S達o Paulo in February 2013 and I have already completed 60\% of my course credits (not counting credits from Texas A\&M) which makes December 2017 my expected graduation date.

\lettersection{Background}
\par My first working experience  was being professor assistant of Introduction to Computer Science class, in University of S達o Paulo. In this opportunity I worked grading students homeworks, answered students questions and also had the opportunity to replace the professor for two of the first lectures. This opportunity led me to my second working experience when the professor I assisted offered me a position as an undergraduate researcher.
\par In my research I worked for Butantan Institute and S達o Paulo Research Foundation researching about the U-Curve problem, an optimization problem that is defined over a boolean lattice and where the objective function describe u-shaped curves for every chain of the lattice. In this opportunity I had to work in a big framework for solving U-Curve problems adding a new algorithm as an attempt to solve the problem more efficiently. While working on this project I had to read papers and existing code; elaborate new code and solutions; and also compare results from new algorithms with the ones that already exists.

\lettersection{Objectives for the Summer}

\par My main goal for the summer is to expand my knowledge in computer science by studying topics related to a research opportunity that I might get. Expanding my knowledge in computer science is an opportunity to get in touch with different areas that may guide me to a specific field that I want to concentrate in.

\par A secondary goal for the summer is to become a more experienced researcher. My first research opportunity was in my university working only with my advisor, therefore doing research in the U.S. is a great opportunity for me to work in a different environment and become a more mature researcher.

\par I also believe that staying in the U.S. for this summer is a great opportunity to meet more academic people. Since I arrived in the United States I met people from a lot of different places and this was great for me to learn from them about different cultures and also to make my contact network wider, which, I believe, can give me more opportunities in the future. If I stay in the U.S. for this summer I will be able to meet even more people and make my contact network even wider.

\end{cvletter}


%-------------------------------------------------------------------------------
% Print the signature and enclosures with above letter informations
%\makeletterclosing

\end{document}
